    \begin{question}{7 (10') (Things Get Real)}
    
    This question is a simplified version of a competitive programming problem, appeared in \textit{CCPC 2022 Final}, where \textit{CCPC} is short for \textit{China Collegiate Programming Contest}. \textbf{You only need to give a solution, without proving it. }

    Little Cyan Fish has devised a new game:
    \begin{itemize}
        \item The game is played with a string $s=s_1 s_2 \cdots s_n$ composed of the digits 0,1 and 2 .
        \item The game proceeds as follows:
        \begin{enumerate}
            \item The player selects an index $i(1 \leq i<n)$ such that $s_i=s_{i+1}$. Notably, if no such $i$ exists, the game ends immediately.
            \item  The characters $s_i$ and $s_{i+1}$ are then removed, and the player returns to step 1 .
        \end{enumerate}
        \item If the player can completely erase the string (leaving an empty string), he wins the game. If not, he loses.
    \end{itemize}

Keen to challenge Little Cyan Fish, Big Flower Letter presents him with a string $s=s_1 s_2 \cdots s_n$ of length $n$. She then proposes $q$ queries of the following type:

\begin{itemize}
    \item $l, r$ : Assuming $t=s[l \cdots r]=s_l s_{l+1} \cdots s_{r-1} s_r$, Little Cyan Fish needs to determine if the player can win the game if they were to play it using the string $t$.
\end{itemize}

Can you help Little Cyan Fish respond to all the queries?

Your algorithm should have  time complexity of $O((n+ q) \log n)$.

\ta{You can assume that, if you can give the right answer to all queries with probability $\ge .9$, you will get \textit{Accepted} in this question. }


Here is an example: 

Let $n = 8, s = 01221220$. 


The answer to query $1, 8$ is \textit{Yes}, since $01221220 \to 011220 \to 0220 \to 00 \to \emptyset$.

The answer to query $2, 5$ is \textit{Yes}, since $1221 \to 11 \to \emptyset$.

The answer to query $3, 7$ is \textit{No}, since $22122 \to 122 \to 1$ is not an empty string. 

The answer to query $5, 8$ is \textit{No}, since $1220 \to 10$ is not an empty string. 
\end{question}