\section*{Problem 2 (15') (Knapsack Counting)} Recall the 0-1 Knapsack problem you have encountered before. We are given $n$ objects of weight $0\leq a_1\leq\cdots\leq a_n\leq b$ where $b$ is the total volume of the knapsack, and we are asked to find a 0-1 assignment $x_1,\cdots,x_n$ such that $\sum_{i=1}^{n}a_ix_i\leq b$. In the counting version, we will compute the number of valid assignments, the set of which is denoted as $S$, 
 instead of finding the optimal solution. The key idea here is similar to $\mathsf{\#DNF}$; that is, sampling from a restricted space. Suppose $b$ is much larger than $n^2$. 
    
    We will use the \textit{scaling method}. Define $a_i':=\lfloor n^2a_i/b\rfloor$, then $0\leq a_i'\leq n^2$. Note the floor operator may cause error when trying to recover $a_i$ from $a_i'$, but the total error is no more than $n(b/n^2)=b/n$. Let $S':=\left\{\bm x: \sum_{i=1}^{n}a_i'x_i\leq n^2\right\}$ be the set of solutions to the rescaled problem. 
    \begin{itemize}
        \item [a. (2')] Prove that $S\subseteq S'$. 
    \end{itemize}
    Apparently, the converse does not necessarily hold. However, we claim that any $\bm x\in S'$ can be shoehorned into $S$ by setting at most one of the $x_i$ from $1$ to $0$. 
    \begin{itemize}
        \item [b. (5')] Prove the claim above, which implies $|S'|\leq (n+1)|S|$. 
    \end{itemize}
    In other words, with probability at least $1/(n+1)$, a uniformly random element of $S'$ is in $S$. So the remaining task here is to compute $|S'|$ and sample from $S'$. Let $C(k,m):=\left|\left\{\bm x: \sum_{i=1}^{k}a_i'x_i\leq m\right\}\right|$. 
    \begin{itemize}
        \item [c. (3')] Design an algorithm based on dynamic programming that computes $C(i,j)$ for all $0\leq i\leq n$ and $0\leq j\leq n^2$. Your algorithm should take $O(n^3)$ time. 
    \end{itemize}
    Definitely, we can fetch the value of $|S'|$ from $C(n,n^2)$. But in fact, we can also use this table to sample uniformly from $S'$. 
    \begin{itemize}
        \item [d. (5')] Based on the values of $C(\cdot,\cdot)$, design an algorithm which outputs a uniform sample of $S'$. Your algorithm should run in $O(n)$ time (excluding the time cost by the random generator).
    \end{itemize}

\begin{answer}
    \begin{enumerate}[label=\alph*).]
        \item For any $\bm x = (x_1, x_2, \cdots, x_n) \in S$, we have 
        \begin{align*}
            \sum_{i=1}^{n} a_i x_i \le b \iff \sum_{i=1}^{n} \frac{n^2 a_i}{b} x_i \le n^2 
        \end{align*}
        And notice that
        \begin{align*}
            \sum_{i=1}^{n} a'_i x_i = \sum_{i=1}^{n} \left\lfloor\frac{n^2 a_i}{b}\right\rfloor x_i \le \sum_{i=1}^{n} \frac{n^2 a_i}{b} x_i \le n^2
        \end{align*}
        Then we have $\bm x \in S'$ which implies $S \subseteq S'$.
        \item Define $k$ be s.t. $ a_j \le b/n$ for $j \le k$ and either $k = n$ or $a_{k+1} > b/n$. Let $C = \{0,1\}^k \times \{0\}^{n-k}$. 
        If $x \in C$ then 
        \begin{align*}
            \sum_{i=1}^{n} a_i x_i \le \sum_{i=1}^{k} a_i x_i \le\frac{bk}{n}  \le n \implies C \subseteq S
        \end{align*}
        For any $\bm x \in S'\setminus S$, then there exists $p = p(x) \in \mathbb{N}^*$, s.t. $x_p = 1$ and $p \notin [k]$ (Otherwise $x \in C \subseteq S$ which is a contradiction).
        Then the sheohorn(or a mapping) from $S'$ to $S$ is defined as:
        \begin{align*}
            f: S' &\to \{0,1\}^n \\
            f(\bm x) &\mapsto \begin{cases}
                \bm x & \text{if } x \in S \\
                \bm y & \text{if } x \in S' \setminus S 
            \end{cases} \\
            &\text{where }  y_j = x_j, \forall j \ne p \text{ and } y_p = 0.
        \end{align*}
        We now prove that for any $\bm x \in S'\setminus S$, $\bm y = f(\bm x) \in S$. Let $\delta_i = a_i - a'_i$, then we have
        \begin{align*}
            \sum_{i=1}^{n} a_i y_i &= \frac{b}{n^2} \sum_{i=1}^{n} (a_i' + \delta_i) y_i = \frac{b}{n^2} \left(\sum_{i=1}^{n} a'_i y_i + \sum_{i=1}^{n}\delta_i y_i \right) \\
            &= \frac{b}{n^2} \left(\sum_{i=1}^{n} a'_i x_i - a'_p + \sum_{i=1}^{n}\delta_i y_i \right) \\
            &\le \frac{b}{n^2} (n^2 - n + n)  = b.
        \end{align*}
        The last inequality holds because $\delta \in [0, 1)$ and $p \notin [k] \implies a_p > b/n \implies a'_p \ge n$. Therefore, $\bm y \in S$ which implies $|S'| \le (n+1)|S|$.

        \item The dynamic programming algorithm is as follows:
        \begin{algo}
            \caption{Dynamic Programming for $|S'|$}
            \label{alg:2.dp}
            \begin{algorithmic}[1]
                \Require{ $a'_1, a'_2, \cdots, a'_n \in [0,n^2]$}
                \Ensure { $C(i,j)$ for all $0\leq i\leq n$ and $0\leq j\leq n^2$.}
                \State $\forall 0\le j \le n^2$, initialize $C(\cdot, \cdot)$ with: 
                \begin{align*}
                    C(1, j) = \begin{cases}
                        2 & \text{if } j \ge a'_1 \\
                        1 & \text{otherwise}
                    \end{cases}
                \end{align*}
                \For {$i = 2$ to $n$} \Comment{Compute the table $C$ using following recursion.}
                    \For {$j = 0$ to $n^2$}
                        \If {$j < a'_i$}
                            \State $C(i,j) = C(i-1,j)$
                        \Else
                            \State $C(i,j) = C(i-1,j) + C(i-1,j-a'_i)$
                        \EndIf
                    \EndFor
                \EndFor
            \end{algorithmic}
        \end{algo}
        The time complexity of the algorithm is $O(n^3)$. 
        \item We can determine a uniform sample from $S'$ by tracing back probabilistically from $C(n,n^2)$ as following:
        \begin{algo}
            \caption{Uniform sampling from $S'$}
            \label{alg:2.sample}
            \begin{algorithmic}[1]
                \Require{ $C(i,j)$ for all $0\leq i\leq n$ and $0\leq j\leq n^2$.}
                \Ensure { A uniform sample from $S'$.}
                \State $j \gets n^2$
                \For{ $i = n$ to $2$}
                    \State $x_i \gets 0$ with prob. $C(i-1,j)/C(i,j)$ else $x_i \gets 1$ with the remaining prob.
                    \If {$x_i = 1$}
                        \State $j \gets j - a'_i$ \Comment{Update the remaining capacity}
                    \Else 
                        \State $j \gets j$ 
                    \EndIf
                \EndFor
                \If{ $C(1, j) = 2$} \Comment{Handle the special case when $i = 1$}
                    \State $x_1 \gets 0$ and $x_1 \gets 1$ with equal prob.
                \Else
                    \State $x_1 \gets 0$ 
                \EndIf
                \State \textbf{Output} $(x_1, x_2, \cdots, x_n)$.
            \end{algorithmic}                
        \end{algo}
        The time complexity of the algorithm is $O(n)$ beacuse each iteration of the for loop takes $O(1)$ time, and the loop runs $n$ times.
    \end{enumerate}
    \ed
\end{answer}